Verification of Stencil Computations through Spatial Specifications

### 2016-07-06 paper.tex

  The predicate is reduced to
  $cons(\textsf{once}(\textsf{exact}(m)),L3,b)$.
%
\item $cons$ predicate from the previous step evaluates to
  $$unique(L3,b) \; \wedge \; cons(\textsf{mult}(\textsf{exact}(m)),L3,b).$$
%
  The former condition holds as no access pattern is repeated in $L3$ by simple
  inspection.
%
\item We further expand the function to 
  \begin{equation*}
    cons(\textsf{mult}(\textsf{up}(m)),L3,b) \wedge 
    cons(\textsf{mult}(\textsf{low}(m),L3,b).
  \end{equation*}
%
\item To evaluate the two predicates above, we need relativised indices for the
  array in question. This is obtained by $analyse$ function from
  Subsection~\ref{subsec:inference}.

  For the example above the relativised indices are as follows:
%
\begin{dmath*}
analyse(L3,b) =
  \{ \vtwo{0}{-1},
     \vtwo{0}{0},
     \vtwo{0}{1},
     \vtwo{-1}{0},
     \vtwo{1}{0}
  \}
\end{dmath*}
%
\item Below we present two match ups of relativised access patterns and unifiers
  to show both the lower and the upper bound holds.

  The unifications for the upper bound:
  \begin{align*}
     \vtwoh{0}{-1} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{0}{0} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{0}{1} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{-1}{0} & \sim_2 \vtwoh{\infty}{0} \\
     \vtwoh{1}{0} & \sim_2 \vtwoh{\infty}{0}
  \end{align*}

  The unifications for the lower bound:
  \begin{align*}
     \vtwoh{-1}{0} & \sim_2 \vtwoh{-1}{\infty} \\
     \vtwoh{0}{0} & \sim_2 \vtwoh{0}{\infty} \\
     \vtwoh{1}{0} & \sim_2 \vtwoh{1}{\infty}  \\
     \vtwoh{0}{-1} & \sim_2 \vtwoh{\infty}{-1} \\
     \vtwoh{0}{0} & \sim_2 \vtwoh{\infty}{0} \\
     \vtwoh{0}{1} & \sim_2 \vtwoh{\infty}{1}
  \end{align*}

  Since both predicates are satisfied the array accesses specified on line three
  are consistent with the \texttt{fivepoint} specification.
\end{enumerate}

Now when $L3$ is slightly changed with an additional term \texttt{b(i+2,j)}, the
new assignment $L3'$ becomes:
%
\begin{minted}[firstnumber=3]{fortran}
a(i,j) = b(i, j-1) + b(i,j) + b(i,j+1) + b(i-1,j) + b(i+1,j) + b(i+2,j+2)
\end{minted}
%
The first four steps will proceed the same way and in the fifth step result
of $\mathit{analyse}(L3',b)$ will have the extra term $\vtwoh{2}{2}$. In sixth
step, since the model has not changed and the relativised indices are only
augmented with a new element, the unification for lower bound still hold.
However, there are not elements in the model, $m$, that can unify with
$\vtwoh{2}{2}$, hence the model fails to hold as an upper bound.

Note that if the specification was as follows:
%
\begin{minted}{fortran}
!= stencil atLeast, readOnce, fivepoint :: b}
\end{minted}
%
The consistency function would have held as
$\mathit{cons}(\textsf{mult}(\textsf{up}(m)),L3',b)$ for the upper bound would
have never been needed. This semantically makes sense since the region
$\texttt{fivepoint}$ is indeed a lower bound for the access pattern of $b$.

>>>>

  The predicate is reduced to
  $cons(\textsf{once}(\textsf{exact}(m)),L3,b)$.
%
\item $cons$ predicate from the previous step evaluates to
  $$unique(L3,b) \; \wedge \; cons(\textsf{mult}(\textsf{exact}(m)),L3,b).$$
%
  The former condition holds as no access pattern is repeated in $L3$ by simple
  inspection.
%
\item We further expand the function to 
  \begin{equation*}
    cons(\textsf{mult}(\textsf{up}(m)),L3,b) \wedge 
    cons(\textsf{mult}(\textsf{low}(m),L3,b).
  \end{equation*}
%
\item To evaluate the two predicates above, we need indexing schemes for the
  array in question. This is obtained by $analyse$ function from
  Subsection~\ref{subsec:inference}.

  For the example above the indexing schemes are as follows:
%
\begin{dmath*}
analyse(L3,b) =
  \{ \vtwo{0}{-1},
     \vtwo{0}{0},
     \vtwo{0}{1},
     \vtwo{-1}{0},
     \vtwo{1}{0}
  \}
\end{dmath*}
%
\item Below we present two sets of generalisations to show both the lower and
  the upper bound holds.

  The generalisations for the upper bound:
  \begin{align*}
     \vtwoh{0}{-1} & \preceq_2 \vtwoh{0}{\infty} \\
     \vtwoh{0}{0} & \preceq_2 \vtwoh{0}{\infty} \\
     \vtwoh{0}{1} & \preceq_2 \vtwoh{0}{\infty} \\
     \vtwoh{-1}{0} & \preceq_2 \vtwoh{\infty}{0} \\
     \vtwoh{1}{0} & \preceq_2 \vtwoh{\infty}{0}
  \end{align*}

  The generalisations for the lower bound:
  \begin{align*}
     \vtwoh{-1}{0} & \preceq_2 \vtwoh{-1}{\infty} \\
     \vtwoh{0}{0} & \preceq_2 \vtwoh{0}{\infty} \\
     \vtwoh{1}{0} & \preceq_2 \vtwoh{1}{\infty}  \\
     \vtwoh{0}{-1} & \preceq_2 \vtwoh{\infty}{-1} \\
     \vtwoh{0}{0} & \preceq_2 \vtwoh{\infty}{0} \\
     \vtwoh{0}{1} & \preceq_2 \vtwoh{\infty}{1}
  \end{align*}

  Since both predicates are satisfied the array accesses specified on line
  three, the stencil comutation is consistent with the \texttt{fivepoint}
  specification.
\end{enumerate}

Now when $L3$ is slightly changed with an additional term \texttt{b(i+2,j+2)},
the new assignment $L3'$ becomes:
%
\begin{minted}[firstnumber=3]{fortran}
a(i,j) = b(i, j-1) + b(i,j) + b(i,j+1) + b(i-1,j) + b(i+1,j) + b(i+2,j+2)
\end{minted}
%
The first four steps will proceed the same way. In the fifth step, result
of $\mathit{analyse}(L3',b)$ will have the extra term $\vtwoh{2}{2}$. In the
sixth step, since the model has not changed and the indexing schemes are only
augmented with a new element, the generalisations for lower bound still hold.
However, there are no elements in the model, $m$, that can generalise
$\vtwoh{2}{2}$, hence the model fails to hold as an upper bound.

Note that if the specification was as follows:
%
\begin{minted}{fortran}
!= stencil atLeast, readOnce, fivepoint :: b
\end{minted}
%
The consistency predicate would have held because for the upper bound
$\mathit{cons}(\textsf{mult}(\textsf{up}(m)),L3',b)$ would not have
been needed. This semantically makes sense since the region
$\texttt{fivepoint}$ is indeed contained in the regions touched by $b$.

\subsection{Synthesis\label{subsec:synthesis}}
